Best practices for forecasting changes in biodiversity: an example using breeding birds

### 2017-08-13 manuscript/draft.Rmd
fix knitting syntax
more abstract

---
title: Best practices for forecasting changes in biodiversity: an example using breeding birds
author:
- David J. Harris
- Shawn D. Taylor
- Ethan P. White
bibliography: refs.bib
csl: csl.csl
geometry: left=3.5cm, right=3.5cm, top=2.25cm, bottom=2.25cm, headheight=12pt, letterpaper
fontsize: 12pt
header-includes:
    - \usepackage{times}
    - \usepackage{setspace}
    - \doublespacing
    - \usepackage{lineno}
    - \linenumbers
output:
  pdf_document: default
  html_document: default
---

@@

Biodiversity forecasts are important for conservation, management, and
evaluating how well current models characterize natural systems. While there are
an increasing number of studies forecasting biodiversity, there is little
information available on how well these forecasts work. Most biodiversity
forecasts are not evaluated in terms of their ability to predict future
diversity, fail to account for uncertainty, and do not use time-series data that
captures the actual dynamics being studied. We compiled a list of ten best
practices for forecasting biodiversity and used hindcasting to evaluate six
different modeling approaches for predicting breeding bird species richness.
Each method was evaluated at `r numbers$N_predict_sites` sites distributed
throughout the continental United States. Forecasts were evaluated for up to a
decade after the final training observations. While all of the models that we
evaluated explained most of the variance in richness, none of them consistently
outperforms a simple baseline model where each site's species richness
fluctuates randomly around a constant value. At present, our best forecast for
future richness is thus to assume that it is unchanging, and current modeling
approaches do not reliably improve on this simple baseline. Most of the best
practices implemented in this study (e.g., evaluating confidence intervals as
well as point estimates) directly influence the outcome of the forecasts; some
even led to large changes in the relative performance of different modeling
approaches. We emphasize the importance of following specific best practices
when making ecological forecasts, and of robust evaluation of model predictions
to facilitate rapid improvement in our biodiversity forecasts.

>>>>

---
title: "Best practices for forecasting changes in biodiversity: an example using breeding birds"
author:
- David J. Harris
- Shawn D. Taylor
- Ethan P. White
bibliography: refs.bib
csl: csl.csl
geometry: left=3.5cm, right=3.5cm, top=2.25cm, bottom=2.25cm, headheight=12pt, letterpaper
fontsize: 12pt
header-includes:
    - \usepackage{times}
    - \usepackage{setspace}
    - \doublespacing
    - \usepackage{lineno}
    - \linenumbers
output:
  pdf_document: default
  html_document: default
---

@@

Biodiversity forecasts are important for conservation, management, and
evaluating how well current models characterize natural systems. While
biodiversity forecasts are common in the literature, there is little information
available on how well these forecasts work. Most biodiversity forecasts are not
evaluated in terms of their ability to predict future diversity, fail to account
for uncertainty, and do not use time-series data that captures the actual
dynamics being studied. We compiled a list of ten best practices for forecasting
biodiversity, and used hindcasting to evaluate six different modeling approaches
for predicting breeding bird species richness. Each method was evaluated at `r
numbers$N_predict_sites` sites distributed throughout the continental United
States. Forecasts were evaluated for up to a decade after the final training
observations. While each model could explain most of the variance in richness,
none of them consistently outperformed a baseline model that simply predicted
constant richness at each site. In particular, we found no evidence that current
methods (such as species distribution models) can turn spatial data into useful
temporal predictions about biodiversity. This result may call into question the
validity of many high-profile forecasts from the last 15 years. Most of the best
practices implemented in this study (e.g., evaluating confidence intervals as
well as point estimates) directly influence the outcome of the forecasts; some
even led to large changes in the relative performance of different modeling
approaches. To facilitate the rapid improvement of biodiversity forecasts, we
emphasize the value of specific best practices in making and evaluating
different forecasting methods.
